Quality Criteria for SKOS Vocabularies

### 2012-03-09 paper.tex
added ld specific criteria/measures

\begin{definition}
Let \(TR \subseteq SR\) be the set of all conceptual resources being transitively related by \texttt{skos:broaderTransitive} respectively \texttt{skos:narrowerTransitive}, not by subproperties thereof. Furthermore, let \(HR \subseteq SR\) be the set of all conceptual resources being related by \textt{skos:broader} and \texttt{skos:narrower} with \((c,c') \in HR \rightarrow (c',c) \in HR\). We can write this criterion as a function \(str:V \rightarrow \mathbb{N}_{0}\) with \[str(v)=\left\|\left\{c \in C : c' \in C, (c,c') \in TR, (c,c') \notin HR\right\}\right\|\]
\end{definition}

\subsection{Linked Data Specific Criteria / Linked Data Measures}

>>>>

\begin{definition}
Let \(TR \subseteq SR\) be the set of all conceptual resources being transitively related by \texttt{skos:broaderTransitive} respectively \texttt{skos:narrowerTransitive}, not by subproperties thereof. Furthermore, let \(HR \subseteq SR\) be the set of all conceptual resources being related by \texttt{skos:broader} and \texttt{skos:narrower} with \((c,c') \in HR \Rightarrow (c',c) \in HR\). We can write this criterion as a function \(str:V \rightarrow \mathbb{N}_{0}\) with \[str(v)=\left\|\left\{c \in C : c' \in C, (c,c') \in TR, (c,c') \notin HR\right\}\right\|\]
\end{definition}

\subsection{Linked Data Specific Criteria / Linked Data Measures}
SKOS has been developed for creating knowledge organization systems that are used ``within the framework of the semantic web" \cite{Bernerslee2001}. Therefore we identified some criteria targeting issues related to best practises in publishing Linked Data, the underlying technology for establishing a Semantic Web \cite{Heath2011}.

\subsubsection{Concept External Link Average} The ``Web of Data" consists of globally interconnected datasets, ``enabling seamless connections between data sets" \cite{Heath2011}. These links can be established between various SKOS vocabularies by ``external links", i.e., references from conceptual resources in a ``source vocabulary" to resources identified by an HTTP URI and available at a different host.

\begin{definition}
Let \(EL=C \times R\) be the set of external links with \(host(c) \neq host(r)\). This criterion can then be written as a function \(elc:V \righarrow \mathbb{ℚ}\) with \[elc(v)=\frac{\left\|EL\right\|}{\left\|C\right\|}\]
\end{definition}

\subsection{HTTP URI Scheme Violation} One of the key concept of Linked Data is that URIs should be dereferencable in order to browse through data in the same way as through ``traditional" Web pages. Using HTTP URIs therefore is the logical consequence and mentioned numerous time in literature \cite{Heath2011}. However, some vocabularies use other URI schemes such as URNs and DOIs when defining resources, which impedes data lookup and crawling purposes.

\begin{definition}
Let \(VR \subseteq C \cup A\) be the set of vocabulary resources identified by an URI (not by a blank node). We can define a function \(httpScheme:R \rightarrow \left\{0,1\right\}\) with \(httpScheme(r)=1\) if the scheme part of the URI is http or https, 0 otherwise. Then this criterion can be written as a function \(usv:V \rightarrow \mathbb{N}_{0}\) with \[usv(v)=\left\|\left\{vr \in VR : httpScheme(vr)=0\right\}\right\|\]
\end{definition}

\subsubsection{Link Target Unavailability} Just as in the ``traditional" Web, resources in the Web of Data might also be unavailable, i.e. resolving their URI leads to an erroneous HTTP response or no response at all. An erroneous HTTP response in that case can be defined as a response code (after a possible redirection) other than 200. Since these ``broken links" hinder information gathering, they should be avoided.

\begin{definition}
Let \(LR \subseteq R\) be the set of all resources being related to conceptual, aggregational and grouping resources. The function \(resp:R \rightarrow \mathbb{N}\) assigns a resource its HTTP response code after following all redirections. We can therefore write this criterion as function \(ltu:V \rightarrow \mathbb{N}_{0}\) with \[ltu(v)=\left\|\left\{r \in LR \cup VR : resp(r)=200\right\}\right\|\]
\end{definition}

\subsubsection{Average Concept In-degree} A periodically updated version of a diagram visualizing ``the data sets in the LOD cloud as well as their interlinkage relationships” is published online \footnote{http://www4.wiwiss.fu-berlin.de/lodcloud/state/}. In these diagrams, DBpedia \footnote{http://dbpedia.org/About} is shown as the most intensely linked resource, i.e. a large number of vocabularies link to resources in that dataset. Therefore a dataset with a large number of other datasets referencing it, can be assumed to be of great value for the community. Accordingly, estimating the number of ``incoming links" of SKOS concepts defined in a vocabulary (henceforth called the in-degree of a concept) would give some hint about how established or accepted a controlled vocabulary is. A method to carry out these estimations would be utilizing existing Linked Data indices like, e.g., Sindice \footnote{http://sindice.com/} that provide a SPARQL endpoint.

\begin{definition}
Let \(U\) be the universe of all datasets on the Web and \(V \in U\). The set \(RR \subseteq R \times R\) contains all resources in distinct datasets related by an RDF property with \((r1,r2) \in RR \Rightarrow host(r1) \neq host(r2)\). This criterion can be written as a function \(aci:V \rightarrow \mathbb{ℚ}\) with \[aci(v)=\frac{\left\|\left\{r \in R : c \in C, (r,c) \in RR\right\}\right|}{\left\|C\right\|}\] 
\end{definition}
